Do motif roles provide unique information about species risk of extinction?

### 2021-06-28 manuscript/motifs_vs_degree.tex

    %\subsection*{1 Why are bottom-up effects important for understanding stability?}
     The plant community is the foundation upon which the myriad of depend for their survival \citep{}. Disturbances in the plant community have been shown to affect important ecosystem properties such as nutrient cycling \citep{} and consumer diversity \citep{scherber2010bottom,Baiser2013}.The stability of the plant community is crucial for sustaining a healthy energy flow in the ecosystem as a whole \citep{Rosenblatt2016} and the structure of the plant community promotes stability across all different levels of ecosystem organization \citep{proulx2010diversity,scherber2010bottom}. Loss of diversity at the basal level typically causes declines in both abundance and richness of all types of consumers, e.g. herbivores, predators, and parasitoids \citep{scherber2010bottom}.
     This has been shown in both theoretical work~\citep{} and empirical experiments ~\citep{} and results from the vertical propagation of disturbances through the network of interacting species.

     %\subsection*{2 Bayesian networks are a fantastic way to model bottom-up effects}
    The study of how an initial decline or extinction leads to loss of other species via direct and indirect effects (i.e., secondary extinctions) is a vibrant area of research \citep{curtsdotter2011robustness, dunne2009cascading, Eklof2006}. Traditionally, there are two main approaches for studying secondary extinctions. First, there are topological models, which are solely based on food web structure \citep{dunne2009cascading}. Here, initial extinctions only affect other species in the network once a consumer has lost all of its prey and therefore must go extinct. The second approach uses dynamical models, which explicitly simulate population dynamics using a system of differential equations \citep{binzer2011susceptibility}. Dynamical models take changes in prey or predator densities into account when calculating the densities of other species in the network. These models therefore describe complex effects such as indirect interactions and are generally more realistic than topological models. However, dynamical models are parameter intensive, can only include a limited number of species, and simulations are time-consuming.

@@

    %\subsection*{3 Global scale network properties are known to affect stability}
    The global structural properties of ecological networks have implications for their functioning \citep{Petchey2002}, stability \citep{Allesina2012} and robustness to disturbances \citep{Dunne2002d, Eklof2006}. For example, more highly connected networks are more resistant to secondary extinctions \citep{Dunne2002d, Eklof2006} and a modular organization where some groups of species are tightly interconnected with limited connections between groups can limit the propagation of disturbances \citep{}.
    %\subsection*{4 meso-scale nw propoerties also affect stability}
    Structural properties on the local scale, i.e. properties of single species, are also well-studied and there is consensus that species with high trophic levels, few prey (low in-degree), and low abundances are generally at higher risks of secondary extinction \citep{binzer2011susceptibility}. However, in-degree and trophic level are both relatively coarse descriptions of a species' role within its food-web~\citep{Cirtwill2018FoodWebs}. 
    For example, specialists may not necessarily have a high risk of extinction if it consumes a basal resource which is at low risk of being affected of disturbance. On the other hand, a consumer with low trophic level may still be at high risk of extinction if it consumes few preys or prey which possess high risk of being affected by disturbance.

@@

    %\subsection{6. only 4 (stable) motifs can appear in Bayesian networks, so how do we expect them to affect stability in a bottom-up context?}
    Of the 13 three-species motifs that can occur in empirical food webs
    \citep{Milo2004}, only four do not include cycles (e.g., A eats B and B eats A or A eats B, B eats C, and C eats A). 
    As cycles must be removed from networks before calculating extinction probabilities in the Bayesian network framework, this study focuses on these four Bayesian-compatible motifs: the three-species chain, omnivory, apparent competition and direct competition motifs. 
    Despite this being a small subset of the motifs which can occur in empirical networks, our choice is justified by studies showing that these four motifs are over-represented in empirical networks and contribute to stability \citep{Stouffer2007, Borrelli2015a, giling2019plant}.   
    When analyzing motifs within a Bayesian Network, it is important to keep in mind that only bottom-up effects are accounted for. Therefore, the implications of the four possible motifs are different than they would be in an empirical network.

    In particular, the lack of top-down effects removes the indirect effects from the two competition motifs. In the case of direct competition, two predators share a common resource. The predators will, within a Bayesian Network, not affect each other indirectly although both will be affected by any disturbance to their resource. Similarly, the two bottom-level species in the apparent competition motif will be independent of each other as there are no top-down effects of the shared consumer to link them.
    The difference between the direct competition and apparent competition motif is thus the number of bottom-level species, accounted for as prey items for a higher level species.
    A predator participating in many direct competition motifs is likely to have fewer prey items than a predator participating frequently in the apparent competition motif, and therefore a higher extinction risk.
    Note, however, that consumers can also appear in these motifs as prey to other consumers.  
    In that case, we would not expect the two motifs to have different effects on extinction risk.

    The structure of the omnivory motif and the three-species chain motif differ from the competition motifs by having an additional trophic level. Therefore, top-level species of these motifs might be more sensitive to disturbances since they can be affected by the loss of the basal resource in the focal motif or an increased probability of extinction of the middle species due to a disturbance to a basal resource not included in the focal motif.

       %\subsection*{Here we test...}
    Here we are interested in how disturbances at the basal level in food webs propagate in the network and whether a consumer's extinction risk depends on 1) the motif profile of the full food web and 2) the motifs in which a focal species participates. In particular, we want to understand whether the information provided by motifs can be explained by the classical global and local food web metrics such as network connectance and species trophic level, or whether motifs provide unique insight into bottom-up extinction.

@@

    \subsection*{How does mean persistence vary with global network structure?}

        Network average persistence of consumers naturally decrease with increasing level of disturbance of basal resources (Table S1, \emph{Appendix SA}). Network size and connectance also affect persistence, but to a much smaller extent (Figure~\ref{fig:lm_CS}). Persistences generally decrease when networks are more connected, but the effect of network size varies depending on level of disturbance: at low disturbance, small and highly-connected webs show the lowest persistence, whereas at high disturbance, large and highly connected webs show the lowest persistence (Fig. S1, \emph{Appendix SA}).

        \begin{figure}[h!]
            \centering
            \includegraphics[width=0.95\textwidth]{figures/persistence_vs_SC_lm.eps}
            \caption{Persistence decreased strongly with increasing probability of disturbance to basal resources, while network size, connectance, and their interaction had smaller effects. \textbf{A)} Persistence decreased slightly with increasing network size at high levels of disturbance, increased slightly with increasing network size at low levels of disturbance, and did not vary noticeably with network size at moderate levels of disturbance. Networks with lower connectance (dotted line) had slightly higher persistence than networks with moderate (solid line) or high (dashed line) connectance.
            \textbf{B)} Persistence decreased with increasing connectance for all levels of disturbance and network size. Network size (line types) did not strongly affect these relationships.}
            \label{fig:lm_CS}
        \end{figure}

    \subsection*{How do network motif profile vary with global network properties?}

        A network's normalized motif profile (i.e., proportions of the total made up by each motif)
        % ; see \emph{Appendix SX} for non-normalized motif profiles) 
        is related to the network's size, connectance, and the interaction between them ($F_{1,2996}$=74.3, $p$\textless0.001; $F_{1,2996}$=4203, $p$\textless0.001; and $F_{1,2996}$=27.8, $p$\textless0.001, respectively in a PERMANOVA of dissimilarity in normalized network motif profiles against network size, connectance, and their interaction).
        Unequal variation in motif profiles across levels of size, connectance, and their interaction may, however, mean that these PERMANOVA results are false positives (Fig. S2, \emph{Appendix SB}).

        The proportions of apparent competition and three-species chains in a network's motif profile tended to decrease with increasing connectance while the proportion of omnivory increased with increasing connectance (Table S2, \emph{Appendix SB}). 
        Network size and the interaction between network size and connectance did not strongly affect the frequencies of most motifs (Fig.~\ref{motif_proportion_lms}). The proportion of direct competition, however, decreased with increasing connectance in small networks but showed very little change in large webs.

@@

    \subsection*{How does mean persistence vary with network motif profile?}

        The motif profile of a network was related to the mean persistence of consumers ($F_{1,2998}$=571, $p$\textless0.001). Mean persistence was higher in networks with higher proportions of the chain, direct competition, and apparent competition motifs (Table S3, \emph{Appendix SC}).Conversely, mean persistence is lower in networks with higher proportions of the omnivory motif. These trends hold for all levels of disturbance (Table S3, \emph{Appendix SC}). Dispersion of motif profiles varies between networks with different levels of persistence ($F_{134,2865}$=2.19$\times10^{24}$, $p$=\textless0.001). For example, networks with higher persistence have more variable motif profiles ($\beta$=0.183, $p$\textless0.001).

@@

    \subsection*{How does species persistence vary with motif participation?}

@@

        \subsubsection*{Interactions between motif participation and disturbance}

@@

        \subsubsection*{Variability of motif effects changes with network properties}

@@

        \begin{figure}
            \centering
            \includegraphics[width=\textwidth]{figures/prop_dens_bp_vs_C_allS.pdf}
            \caption{Density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species. The slopes are derived from standard linear regression. Columns show the result for various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (left) to $\pi_{disturbed} = 0.5$ (right). Rows show various levels of connectance. The dotted, vertical line indicate zero on the x-axis. A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increase persistence. The fraction of replicates with a slope greater than zero are stated in numbers in each sub-plot, the color corresponding to each motif (legend). }
            \label{fig:density_prop_C}
        \end{figure}

        \begin{figure}[h!]
             \centering
             \includegraphics[width=\textwidth]{figures/prop_dens_bp_vs_S_allC.pdf}
            \caption{Density (y-axis) of slopes (x-axis) for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species. The slopes are derived from standard linear regression. A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increase persistence. The fraction of replicates with a slope greater than zero are stated in numbers in each sub-plot, the color corresponding to motif (legend). Columns show the result for various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (left) to $\pi_{disturbed} = 0.5$ (right). Rows show various levels of network size. The vertical line indicate zero on the x-axis.}
            \label{fig:density_prop_S}
        \end{figure}

    \subsection*{How does persistence vary with degree and trophic level?}

@@

    \subsection*{Do simpler measures explain effects of motifs?}

        \subsubsection*{Motif participation and network size and connectance}

        Motif participation varied with web size, connectance, and their interaction (Fig. S4, Table S6; \emph{Appendix SG}).
        Species in all networks appeared most frequently in the apparent competition motif, but the frequency of the apparent competition motif decreased with increasing network size and connectance.
        The frequency of the omnivory motif increased sharply with increasing connectance, while the frequency of the direct competition motif increased slightly with increasing network size. 
        Species' motif roles also vary with their in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}).
        Species with higher in-degree tend to have higher proportions of their motif role made up by omnivory. Proportions of the other three motifs tend to decline with increasing degree.

        Species with higher trophic levels tend to have higher proportions of apparent competition and three-species chains.

@@

    \subsection*{How does persistence vary with motif positions?}

        A species probability of persisting is significantly related to the frequency of all positions, disturbance, and the interaction between frequency of each position and disturbance.
        \begin{table}[h!]
            \caption{Coefficients in a series of regressions relating probability of persistence to the frequency of a position in a species' motif role, disturbance, and their interaction. For ease of comparison between motifs, we did not center or scale predictors. Note, however, that some motifs and positions have higher frequencies than others.}
            \label{tab:persistence_vs_positions}
            \centering
            \begin{tabular}{l | c c c c}
            Position	& Intercept	& Disturbance	& Position	& Interaction \\
            \hline
            Chain: top	& 0.880 &	-0.985 & -0.00544 & -0.332 \\
            Chain: bottom	& 0.883 &	-1.07 &	-0.0510 &	0.647 \\
            Chain: middle	& 0.862 & -1.00 &	0.246 &	-0.244 \\
            \hline
            Dir. Comp.: top & 0.872 &	-0.995	& 0.0549 &	-0.178 \\
            Dir. Comp.: bottom & 0.886 &	-1.07 &	-0.113 &	0.939 \\
            \hline
            Omnivory: top	& 0.872 &	-0.984 &	0.180 &	-0.818 \\
            Omnivory: middle & 0.873 & -1.00 &	0.119 &	-0.324 \\
            Omnivory: bottom & 0.882 & -1.06 &	-0.0449 &	0.845 \\
            \hline
            App. Comp.: top & 0.865 &	-0.966 &	0.140 &	-0.520 \\
            App. Comp.: bottom & 0.910 &	-1.12 &	-0.100 &	0.323 \\
            \hline
            \end{tabular}
        \end{table}

        % Organizing results by motif - keep this OR by no. prey
        In the omnivory motif, persistence increased with the frequency of the bottom position (two predators, no prey within the motif) across all levels of disturbance but most steeply at high levels of disturbance (Fig.~\ref{figs:persistence_position_motifs}).
        Peristence increased slightly with the frequency of the middle position (one prey, one predator) at low levels of disturbance but showed little relationship at higher levels of disturbance.
        Persistence increased with the frequency of the top position (two prey, no predators within the motif) at low disturbance but decreased at higher levels of disturbance.

@@

        \begin{figure}[h!]
        \includegraphics[width=.95\textwidth]{figures/persistence_positions_bymotif.eps}
        \label{figs:persistence_position_motifs}
        \caption{A species' probability of persistence varied with the frequency of different motif positions in its role. Relationships are shown for the range 0-0.8 for all positions, but some positions (especially in the omnivory motif) are less common. Line colour indicates the probability of extinction for basal resources. Line style indicates the number of prey within the motif for each position. Symbol shape indicates the number of predators within the motif for each position; note that only bottom-up extinctions were included in our simulations.}
        \end{figure}

        % Organizing by number of prey
        Across the four motifs we consider, positions with no prey within the motif tended to show weak relationships to persistence at low levels of disturbance and be associated with increased persistence at high levels of disturbance (Fig.~\ref{fig:persistence_positions_prey}).
        These trends were strongest for direct competition and omnivory and weakest for apparent competition.
        The two positions with stronger effects indicate a species with two predators within the motif, while the weaker effects are for motifs with one predator within the motif.

        Positions with one prey within the motif were more variable.
        The top position in the direct competition motif and middle position in the omnivory motif (both of which have one prey) were associated with increased persistence at low levels of disturbance and lower persistence at higher levels of disturbance, but these effects were relatively weak.
        The top and middle positions of the three-species chain both include a single prey; the middle position was associated with increased persistence at all levels of disturbance while the top position was associated with decreased persistence at all levels of disturbance.

        Positions with two prey in the apparent competition and omnivory motifs showed very similar patterns.
        At low levels of disturbance, persistence increased with the frequency of either position.
        At high levels of disturbance, persistence decreased as the frequency of either motif increased.
        This echoes the pattern for in-degree.

        \begin{figure}
            \centering
            \includegraphics[width=.95\textwidth]{figures/persistence_positions_byprey.eps}
            \caption{A species' probability of persistence varied with the frequency of different motif positions in its role. Relationships are shown for the range 0-0.8 for all positions, but some positions (especially in the omnivory motif) are less common. Line colour indicates the motif containing a position. Line style indicates the number of prey within the motif for each position. Symbol shape indicates the number of predators within the motif for each position; note that only bottom-up extinctions were included in our simulations. The top set of lines corresponds to a probability of extinction for basal resources of 0.1. The middle set of lines corresponds to a probability of extinction for basal resources of 0.26, and the lowest set of lines corresponds to a probability of extinction for basal resources of 0.5.}
            \label{fig:persistence_positions_prey}
        \end{figure}

    \subsection*{How do motif positions vary with other network properties?}

        \subsubsection*{Degree and trophic level}

            The top position in the omnivory motif increased with increasing degree across all trophic levels.
            The frequency of the middle position increased with the interaction of degree and trophic level.
            The bottom position was generaly rare, especially for species with high degree or high trophic level (this suggests that the bottom position is frequently a basal resource).

            The frequency of the bottom position in the apparent competition motif decreased strongly with increasing degree, especially for species at a high trophic level.
            The frequency of the top position increased strongly with increasing degree and was not strongly related to trophic level.

            The frequency of the bottom position in the direct competition motif increased with increasing degree, especially for species at a high trophic level.
            The bottom position was very rare, especially for species with high degree (also suggesting that this position is usually a basal resource).

            The bottom position in the three-species chain was generally rare, especially for species with high degree or trophic level (again, this is likely to commonly be basal resources).
            The middle position in the three-species chain did not vary strongly with degree and declined slightly with increasing trophic level.
            The top position in the three-species chain increased with trophic level, degree, and their interaction.

        \subsubsection*{Network size and connectance}

            % Many of these are significant, but the slopes tend to be very shallow. TL and degree are more interesting - this is DEFINITELY SI.
            The frequency of the top position in the apparent competition motif did not vary strongly with network size or connectance.
            The frequency of the bottom position in the apparent competition motif decreased with increasing connectance but did not vary strongly vary with network size.

            The frequency of the top position in the direct competition motif increased slightly with the interaction of network size and connectance (i.e., in large and highly-connected networks).
            The frequency of the bottom position in the direct competition motif was low in all networks.

            The frequency of all three positions in the omnivory motif showed similar relationships with network size and connectance.
            Frequencies of these motifs increased with increasing connectance and were not strongly affected by network size.
            None of the positions in the three-species chain varied strongly with network size or connectance.

@@

    [[Big deal result: connectance is important for motif structure of network, size not so much]]

    [[Persistence depends on S, C, interaction with $basal_p$ (lowest persistence swaps from highest S to lowest C in Figs. 1-2, high C in both cases - but small absolute differences so maybe we don't want to make too big a deal out of this)]]

    [[proportion of omnivory increases with increasing C, while the other doesn't. Connect to later findings involving omnivory]]

    \subsection*{How does persistence vary with motif positions?}

        A species probability of persisting is significantly related to the frequency of all positions, disturbance, and the interaction between frequency of each position and disturbance (Table S7, \emph{Appendix SH}).

        % Organizing results by motif - keep this OR by no. prey
        In the omnivory motif, persistence increased with the frequency of the bottom position (two predators, no prey within the motif) across all levels of disturbance but most steeply at high levels of disturbance (Fig.~\ref{figs:persistence_position_motifs}).
        Peristence increased slightly with the frequency of the middle position (one prey, one predator) at low levels of disturbance but showed little relationship at higher levels of disturbance.
        Persistence increased with the frequency of the top position (two prey, no predators within the motif) at low disturbance but decreased at higher levels of disturbance.

        \begin{figure}[h!]
        \includegraphics[width=.95\textwidth]{figures/persistence_positions_bymotif.eps}
        \label{figs:persistence_position_motifs}
        \caption{A species' probability of persistence varied with the frequency of different motif positions in its role. Relationships are shown for the range 0-0.8 for all positions, but some positions (especially in the omnivory motif) are less common. Line colour indicates the probability of extinction for basal resources. Line style indicates the number of prey within the motif for each position. Symbol shape indicates the number of predators within the motif for each position; note that only bottom-up extinctions were included in our simulations.}
        \end{figure}

        % Organizing by number of prey
        Across the four motifs we consider, positions with no prey within the motif tended to show weak relationships to persistence at low levels of disturbance and be associated with increased persistence at high levels of disturbance (Fig.~\ref{fig:persistence_positions_prey}).
        These trends were strongest for direct competition and omnivory and weakest for apparent competition.
        The two positions with stronger effects indicate a species with two predators within the motif, while the weaker effects are for motifs with one predator within the motif.

        Positions with one prey within the motif were more variable.
        The top position in the direct competition motif and middle position in the omnivory motif (both of which have one prey) were associated with increased persistence at low levels of disturbance and lower persistence at higher levels of disturbance, but these effects were relatively weak.
        The top and middle positions of the three-species chain both include a single prey; the middle position was associated with increased persistence at all levels of disturbance while the top position was associated with decreased persistence at all levels of disturbance.

        Positions with two prey in the apparent competition and omnivory motifs showed very similar patterns.
        At low levels of disturbance, persistence increased with the frequency of either position.
        At high levels of disturbance, persistence decreased as the frequency of either motif increased.
        This echoes the pattern for in-degree.

        \begin{figure}
            \centering
            \includegraphics[width=.95\textwidth]{figures/persistence_positions_byprey.eps}
            \caption{A species' probability of persistence varied with the frequency of different motif positions in its role. Relationships are shown for the range 0-0.8 for all positions, but some positions (especially in the omnivory motif) are less common. Line colour indicates the motif containing a position. Line style indicates the number of prey within the motif for each position. Symbol shape indicates the number of predators within the motif for each position; note that only bottom-up extinctions were included in our simulations. The top set of lines corresponds to a probability of extinction for basal resources of 0.1. The middle set of lines corresponds to a probability of extinction for basal resources of 0.26, and the lowest set of lines corresponds to a probability of extinction for basal resources of 0.5.}
            \label{fig:persistence_positions_prey}
        \end{figure}

    \subsection*{How do motif positions vary with other network properties?}

        The frequencies of motif positions in a species' role varied strongly and significantly with degree and trophic level, but only weakly with network size and connectance (\emph{Appendix SI}).

           \subsubsection*{Degree and trophic level}

>>>>

    %\subsection*{1 Why are bottom-up effects important for understanding stability?}
     The plant community is the foundation upon which the myriad of species in food webs depend for their survival \citep{}. Disturbances in the plant community have been shown to affect important ecosystem properties such as primary \citep{} and secondary production \citep{}, soil respiration and carbon cycling \citep{} and consumer diversity \citep{scherber2010bottom,Baiser2013}. The stability of the plant community is crucial for sustaining a healthy energy flow in the ecosystem as a whole \citep{Rosenblatt2016} and the structure of the plant community promotes stability across all different levels of ecosystem organization \citep{proulx2010diversity,scherber2010bottom}. Loss of diversity at the basal level typically causes declines in both abundance and richness of all types of consumers, e.g. herbivores, predators, and parasitoids \citep{scherber2010bottom}.
     %This has been shown in both theoretical work~\citep{} and empirical experiments ~\citep{} and results from the vertical propagation of disturbances through the network of interacting species.

     %\subsection*{2 Bayesian networks are a fantastic way to model bottom-up effects}
    The study of how an initial population decline or extinction leads to loss of other species via direct and indirect effects (i.e., secondary extinctions) is a vibrant area of research \citep{curtsdotter2011robustness, dunne2009cascading, Eklof2006}. Traditionally, there are two main approaches for studying secondary extinctions. First, there are topological models, which are solely based on food web structure \citep{dunne2009cascading}. Here, initial extinctions only affect other species in the network once a consumer has lost all of its prey and therefore must go extinct. The second approach uses dynamical models, which explicitly simulate population dynamics using a system of differential equations \citep{binzer2011susceptibility}. Dynamical models take changes in prey or predator densities into account when calculating the densities of other species in the network. These models therefore describe complex effects such as indirect interactions and are generally more realistic than topological models. However, dynamical models are parameter intensive, can only include a limited number of species, and simulations are time-consuming.

@@

    %\subsection*{3 Global scale network properties are known to affect stability}
    The global structural properties of ecological networks have implications for their functioning \citep{Petchey2002}, stability \citep{Allesina2012} and robustness to disturbances \citep{Dunne2002d, Eklof2006}. For example, more highly connected networks are more resistant to secondary extinctions \citep{Dunne2002d, Eklof2006} and a modular organization, where some groups of species are tightly interconnected with limited connections between groups, can limit the propagation of disturbances \citep{}.
    %\subsection*{4 meso-scale nw propoerties also affect stability}
    Structural properties on the local scale, i.e. properties of single species, are also well-studied and there is consensus that species with high trophic levels, few prey (low in-degree), and low abundances are generally at higher risks of secondary extinction \citep{binzer2011susceptibility}. However, in-degree and trophic level are both relatively coarse descriptions of a species' role within its food-web~\citep{Cirtwill2018FoodWebs}. 
    For example, specialists may not necessarily have a high risk of extinction if it consumes a basal resource which is at low risk of being affected of disturbance. On the other hand, a consumer with low trophic level may still be at high risk of extinction if it consumes preys which possess high risk of being affected by disturbance.

@@

    %\subsection{6. only 4 (stable) motifs can appear in Bayesian networks, so how do we expect them to affect stability in a bottom-up context?}
    Of the 13 three-species motifs that can occur in empirical food webs there are four motifs that are over-represented in empirical networks \citep{Stouffer2007} and contribute to stability \citep{Stouffer2007, Borrelli2015a, giling2019plant}: the three-species chain, omnivory, apparent competition and direct competition motifs.
    These four motifs are also the only ones which can occur in the acyclic networks required for Bayesian network simulations~\citep{Eklof2013}. Because of this restriction, we only consider these four particularly important motifs.  
    % only four do not include cycles (e.g., A eats B and B eats A or A eats B, B eats C, and C eats A). 
    %As cycles must be removed from networks before calculating extinction probabilities in the Bayesian network framework, this study focuses on these four Bayesian-compatible motifs:

    %When analyzing motifs within a Bayesian Network, it is important to keep in mind that only bottom-up effects are accounted for. Therefore, the implications of the four possible motifs are different than they would be in an empirical network.

    %In particular, the lack of top-down effects removes the indirect effects from the two competition motifs. In the case of direct competition, two predators share a common resource. The predators will, within a Bayesian Network, not affect each other indirectly although both will be affected by any disturbance to their resource. Similarly, the two bottom-level species in the apparent competition motif will be independent of each other as there are no top-down effects of the shared consumer to link them.
    %The difference between the direct competition and apparent competition motif is thus the number of bottom-level species, accounted for as prey items for a higher level species.
   % A predator participating in many direct competition motifs is likely to have fewer prey items than a predator participating frequently in the apparent competition motif, and therefore a higher extinction risk.Note, however, that consumers can also appear in these motifs as prey to other consumers.  In that case, we would not expect the two motifs to have different effects on extinction risk.
    %The structure of the omnivory motif and the three-species chain motif differ from the competition motifs by having an additional trophic level. Therefore, top-level species of these motifs might be more sensitive to disturbances since they can be affected by the loss of the basal resource in the focal motif or an increased probability of extinction of the middle species due to a disturbance to a basal resource not included in the focal motif.

       %\subsection*{Here we test...}
    Here we are interested in how disturbances at the basal level in food webs propagate up the network and whether a consumer's extinction risk depends on 1) the motif profile of the full food web and 2) the motifs in which a focal species participates. In particular, we want to understand whether the information provided by motifs can be explained by the classical global and local food web properties such as network connectance and species trophic level, or whether motifs provide unique insight into bottom-up extinction cascades.

@@

    \subsection*{Persistence did not vary strongly with global network structure}

        Network average persistence of consumers naturally decreases with increasing level of disturbance of basal resources (Table S1, \emph{Appendix SA}). Network size and connectance also affect persistence, but to a much smaller extent (Figure S1, \emph{Appendix SA}). Persistences generally decrease when networks are more connected, but the effect of network size varies depending on level of disturbance: at low disturbance, small and highly-connected webs show the lowest persistence, whereas at high disturbance, large and highly connected webs show the lowest persistence (Fig. S2, \emph{Appendix SA}).

    \subsection*{Network motif profiles varied with global network properties}

        A PERMANOVA test suggests that a network's normalized motif profile as a whole (i.e., proportions of the total made up by each motif) is related to the network's size, connectance, and the interaction between them, but unequal variation in motif profiles across global network properties may, however, mean that these PERMANOVA results are false positives (Fig. S2, \emph{Appendix SB}).
        Considering each motif separately, the proportions of apparent competition and three-species chains in a network's motif profile tended to decrease with increasing connectance while the proportion of omnivory increased with increasing connectance (Table S2, \emph{Appendix SB}). 
        Network size and the interaction between network size and connectance did not strongly affect the frequencies of most motifs (Fig.~\ref{motif_proportion_lms}). The proportion of direct competition, however, decreased with increasing connectance in small networks but showed very little change in large webs.

@@

    \subsection*{Persistence decreases with greater omnivory in a network}

        The motif profile of a network was related to the mean persistence of consumers ($F_{1,2998}$=571, $p$\textless0.001). Mean persistence was higher in networks with higher proportions of the chain, direct competition, and apparent competition motifs (Table S3, \emph{Appendix SC}). Conversely, mean persistence is lower in networks with higher proportions of the omnivory motif. These trends hold for all levels of disturbance (Table S3, \emph{Appendix SC}). Dispersion of motif profiles varies between networks with different levels of persistence ($F_{134,2865}$=2.19$\times10^{24}$, $p$=\textless0.001). For example, networks with more variable motif profiles have higher persistence ($\beta$=0.183, $p$\textless0.001).

@@

    \subsection*{Species persistence varies with motif participation}

@@

        \subsubsection*{Effects of motif participation depend on disturbance}

@@

        \subsubsection*{Effects of motif participation are more consistent in highly-connected networks}

@@

        \begin{figure}[h!]
            \centering
            \includegraphics[width=\textwidth]{figures/prop_dens_bp_vs_C_allS.pdf}
            \caption{Density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species. The slopes are derived from standard linear regression. Columns show the result for various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (left) to $\pi_{disturbed} = 0.5$ (right). Rows show various levels of connectance. The dotted, vertical line indicate zero on the x-axis. A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increase persistence. The fraction of replicates with a slope greater than zero are stated in numbers in each sub-plot, the color corresponding to each motif (legend). }
            \label{fig:density_prop_C}
        \end{figure}

    \subsection*{Persistence increases with degree but decreases with STL}

@@

    \subsection*{Simpler measures do not fully explain effects of motifs}

        Motif participation varied with web size, connectance, and their interaction (Fig. S4, Table S6; \emph{Appendix SG}).
        Species in all networks appeared most frequently in the apparent competition motif, but the frequency of the apparent competition motif decreased with increasing network size and connectance.
        The frequency of the omnivory motif increased sharply with increasing connectance, while the frequency of the direct competition motif increased slightly with increasing network size.

        Species' motif roles also vary with their in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}).
        Species with higher in-degree tend to have higher proportions of their motif role made up by omnivory. Proportions of the other three motifs tend to decline with increasing degree.
        Species with higher trophic levels tend to have higher proportions of apparent competition and three-species chains.

@@

    \subsection*{Persistence varies with a species' motif role}

        A species probability of persisting is significantly related to the frequency of all positions, disturbance, and the interaction between frequency of each position and disturbance (Table SX, \emph{Appendix SH}).
        In the omnivory motif, persistence increased with the frequency of the bottom position (two predators, no prey within the motif) across all levels of disturbance but most steeply at high levels of disturbance (Fig.~\ref{figs:persistence_position_motifs}).
        Peristence increased slightly with the frequency of the middle position (one prey, one predator) at low levels of disturbance but showed little relationship at higher levels of disturbance.
        Persistence increased with the frequency of the top position (two prey, no predators within the motif) at low disturbance but decreased at higher levels of disturbance.

@@

    \begin{figure}[h!]
        \includegraphics[width=.95\textwidth]{figures/persistence_positions_bymotif.eps}
        \caption{A species' probability of persistence varied with the frequency of different motif positions in its role. Relationships are shown for the range 0-0.8 for all positions, but some positions (especially in the omnivory motif) are less common. Line colour indicates the probability of extinction for basal resources. Line style indicates the number of prey within the motif for each position. Symbol shape indicates the number of predators within the motif for each position; note that only bottom-up extinctions were included in our simulations.}
        \label{figs:persistence_position_motifs}
        \end{figure}

        The frequencies of motif positions in a species' role varied with network properties (\emph{Appendix SI}).
        Motif roles had stronger relationships with degree and STL than with network size or connectance.

@@

Somewhere early in discusssion (needs tidying up):

Apparent comp and direct comp have no indirect effects in BN, so we should be able to explain them with simpler properties.
AC is strongly negatively correlated with in-degree. Higher in degree is associated with both high persistence at low disturbance (Fig. 7A) and low participation in AC (Fig. 7C). Low participation in AC is in turn associated with high persistence at low disturbance (Fig. 4). Note that trends for degree and AC reverse at high disturbance in the same way!
DC is not strongly correlated with degree but does decrease strongly at high trophic levels. High trophic level and low DC are both associated with low persistence at all levels of disturbance, making it likely that the effects we observe for DC are trophic level in disguise.

Three-species chain and omnivory do have indirect effects in a BN setting, and they can't be fully explained by degree or trophic level (since the result doesn't follow the result for degree). We can see this because despite the strong correlations with degree and trophic level these motifs don't show the same patterns with persistence. We therefore focus on them here.

\section*{Discussion}
% Connect back to intro -- importance of the plant community for a healthy consumer community.
% Focus on what additional information is revealed from analyzing motif compared to global network properties or species characteristics.
% Connect to traits -- Cirtwill & Eklöf, 2018, Ecol Lett. 
% Bayesian network approach - what extra information does motifs provide when we are interested in bottom- up effects. 
% * Difference between 3-sp chain and omnivory, the importance of the rescue prey in omnivory (in dynamical sims omni was mostly bad, 3-sp chain good). Omni link destabilizing if strong, stabilizing if weak (other studies). In BN just a link as important as other links
% * Omnivory in fig 5 - high connectance, always good, not as good with lower levels of C
% * Extra prey in omnivory changes from being a security to a liability (with increasing disturbance?)
% * ADD REFS - high disturbance scenario is analogous to decreasing plant diversity

% Results summary - network level (following Anna)
Changes in the primary producer level in food webs have profound effects on all other species in the network \citep{}. In particular, the disturbances can percolate across trophic levels as all consumers depend directly or indirectly on primary producers for their survival \citep{}. Understanding which characteristics of a network and its species influence the severity of outcomes after a disturbance is crucial, e.g. for conservation actions \citep{}. 
Here we were particularly interested in how the distribution of meso-scale structures in ecological networks are related to average network persistence when the basal level is disturbed by species loss. In order to disentangle the effects caused by the motif distribution we also analyzed global (network-wide) and local (species) properties.
%Consistent with earlier work, networks with higher connectance or species richness tend to have lower persistences~\citep{}.
A network's motif profile had relatively strong relationships to the network mean persistence, while effects from global scale structures, such as network size and connectance, were relatively small. 
However, connectance (and, to a lesser extent network size) do affect network motif profiles. Thus, global-scale properties seem to have stronger effects on network persistence by changing meso-scale properties than directly. ~\citep{prill2005dynamic, bascompte2005simple}.

%In all cases, the effect of increasing disturbance to basal resources had a much larger effect on consumer persistences than did network structure. This is a consequence of the bottom-up calculations done to receive the final persistence. The importance of the basal level for network robustness has been demonstrated in numerous studies \citep{scherber2010bottom}, as well as the sensitivity of high trophic levels (REF).

% Network-level motif profiles, persistence, and relationships to S and C.
A network's mean persistence depended on which motifs were most common in the network profile. In general, persistence was lowest in networks with high frequencies of the omnivory motif and highest in networks with high frequencies of the apparent competition motif. Direct competition and three-species chain motifs had smaller effects. The effects were clearly pronounced when disturbance was low but attenuated with increased disturbance.
The negative relationship between omnivory and mean persistence is consistent with other studies showing that an omnivory motif is less likely to be stable than the other motifs considered~\citep{Borrelli2015a}.

High frequency of the apparent competition motif was correlated with an on average higher in-degree for consumers. In-degree in turn is often related to higher persistence \citep{}.
